### 2016-03-17 Draft.tex
Add commments about organization of the manuscript

@@

@@

\begin{figure}[h!]\centering
\includegraphics[width=0.68\textwidth]{Hazards.pdf}
\caption{\label{HazardFxns}  Examples of instantaneous rates of detection (hazard functions) from exponential, gamma, lognormal, and Weibull TTDDs.}
\end{figure}

@@

Our model is formulated for single-visit removal-sampled point-count surveys conducted at a set of sites $s=1,\dotso,S$ repeated across distinct time periods $r=1,\dotso,R$.  We present the model for times to detection censored into $i=1,\dotso,I$ intervals, but it is straight-forward to simplify the model for exact time-to-detection data.  Our data therefore take the form of counts indexed by site, time period, and interval: $n_{sri}$.  For any given survey (uniquely defined by site and time period), we write the vector of counts across all $I$ intervals as $\textbf{n}_{sr}$.

@@

\subsubsection{Data Model}
We assume that counts from different surveys are conditionally independent.  We also assume that the time-to-detection is independent, identically distributed for all individuals of a species at any given survey.  Given that an organism is detected, we denote the probability it is first detected during interval $i$ as $p_{sri}$ such that $\sum_{i=1}^I p_{sri} = 1$.  We write $\textbf{p}_{sr}$ to denote the vector of interval-specific detection probabilities at any given survey.  The data model is then:

\begin{align}\label{eq:datamodel}
\textbf{n}_{sr} &\sim \mbox{Multinomial}\left(n_{sr}^{(obs)}, \textbf{p}_{sr}\right)\\
n_{sr}^{(obs)} &\sim \mbox{Binomial}\left(N_{sr}, p_{sr}^{(det)}\right)
\end{align}

@@

\subsubsection{Detection Model}\label{sec:detectionmodel}
We model $\textbf{p}_{sr}$ and $p_{sr}^{(det)}$ by means of the partitioned cdf of a TTDD, which we express generically as $F_T(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})$.  Here, $T$ is a random variable for the first time at which an individual is detected, $\textbf{X}_{sr}^D$ and $\textbf{Z}_{sr}^D$ are detection-related covariates and random effect factor levels to be used in a mixed effects linear regression, and $\boldsymbol{\theta}$ is a vector of parameters appropriate to the TTDD being used.  In this study, we consider eight different forms for the TTDD.  The first four TTDDs we employ are:
\begin{enumerate}
\item $T_{srq} \overset{ind}{\sim} \text{Exponential}(\varphi_{sr})$
\item $T_{srq} \overset{ind}{\sim} \text{Gamma}(\alpha, \varphi_{sr})$
\item $T_{srq} \overset{ind}{\sim} \text{Weibull}(k, \varphi_{sr})$
\item $T_{srq} \overset{ind}{\sim} \text{Lognormal}(\mu_{sr}, \sigma_{det}^2)$
\end{enumerate}
where $q = 1,\dotso,N_{sr}$ indexes individuals present at survey $sr$, and $\varphi_{sr}$ is always a survey-specific detection rate parameter.  We treat the shape parameters $\alpha$, $\sigma_{det}^2$, and $k$ as constant across all surveys.

@@

For exponential, gamma, and Weibull TTDDs, we model the log of the rate parameter $\varphi_{sr}$ via mixed effects regression using survey-specific covariates.  In so doing, we are equivalently modeling the log of the expected time-to-detection, which is inversely related to the detection rate.  For the lognormal distribution, we fit a regression to $-\mu_{sr}$.  Thus, the link functions in our detection model are:
\begin{alignat}{3}
&\text{Exponential:\;} &&\log(\varphi_{sr}) &&= \text{Intercept}^D + \textbf{X}_{sr}^D\boldsymbol{\beta}^D + \textbf{Z}_{sr}^D\boldsymbol{\xi}^D = -\log(E[T_{srq}])\\
&\text{Gamma:} &&\log(\varphi_{sr}) &&= \text{Intercept}^D + \textbf{X}_{sr}^D\boldsymbol{\beta}^D + \textbf{Z}_{sr}^D\boldsymbol{\xi}^D = -\log(E[T_{srq}]) + \log(\alpha)\\
&\text{Weibull:}  &&\log(\varphi_{sr}) &&= \text{Intercept}^D + \textbf{X}_{sr}^D\boldsymbol{\beta}^D + \textbf{Z}_{sr}^D\boldsymbol{\xi}^D = -\log(E[T_{srq}]) + \log\left(\Gamma\left(1 + 1/k\right)\right)\\
&\text{LogNormal:} &&-\mu_{sr} &&= \text{Intercept}^D + \textbf{X}_{sr}^D\boldsymbol{\beta}^D + \textbf{Z}_{sr}^D\boldsymbol{\xi}^D = -\log(E[T_{srq}]) + \sigma_{det}^2/2
\end{alignat}
where $\boldsymbol{\beta}^D$ is a vector of fixed effects, and $\xi_i^D \overset{ind}{\sim} N(0,\sigma_{D[i]}^2)$ are random effects, with $\sigma_{D[i]}^2$ denoting the variance associated with $\xi_i^D$.  When the TTDD is a mixture distribution, we model $F_{T1}$ portion of the mixture according to the above link functions and treat the mixing parameter $\gamma$ as constant across all surveys.

For interval-censored data, we denote the cutoffs for each observation interval with $C_i$, where $C_0 = 0 \text{ and } C_I = \text{the survey duration}$.  Conditioned on an individual being detected, we calculate interval-specific detection probabilities $p_{sri}$ by partitioning the cdf of the TTDD at the observation interval cutoffs and normalizing by the binomial detection probability $p_{sr}^{(det)}$.  In summary, the detection model is:
\begin{align}
p_{sri} &= \left[F_T(C_i|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta}) - F_T(C_{i-1}|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})\right]\big/p_{sr}^{(det)}\\
\label{eq:pdet}p_{sr}^{(det)} &= F_T(C_I|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})\\
F_T(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta}) &= \text{a TTDD as described above}
\end{align}

To adapt our model for exact time-to-detection data, we only need to replace the multinomial distribution in Eq. (\ref{eq:datamodel}).  Instead of modeling the interval-specific counts $\textbf{n}_{sr}$, we model the times to detection directly.  Defining $\textbf{T}_{sr}$ as a vector of all observed detection times for a given survey, and continuing to treat all detections as independent, we write
\[
Pr\left(\textbf{T}_{sr}|n_{sr}^{(obs)}\right) = \prod\limits_{q=1}^{n_{sr}^{(obs)}} f_T(t_{srq}|t_{srq}<C_I)
\]
The rest of the model remains the same, except there is no need to calculate interval-specific times to detection.  Partial finite-mixture TTDDs clearly only make sense in the interval-censored case.

\subsubsection{Abundance Model}
For simplicity, we model abundance at each site during each year $N_{sr}$ using a Poisson distribution with a survey-specific expected value of $\lambda_{sr}$, though any non-negative integer-valued distribution may be used.  We can decompose $N_{sr}$ into observed and unobserved portions.  The unobserved portion follows a binomial-Poisson hierarchy, resulting in Equation \ref{eq:binPois}.  We model expected abundance $\lambda_{sr}$ using a mixed effects linear regression with a log link.

@@

@@

@@

@@

\subsubsection{Priors}\label{sec:priors}
Our goal in selecting priors is to be non-informative within a reasonable range of values.  Priors for Intercept$^D$ are centered so that the median detection probability is $p_{sr}^{(det)} = 0.50$, based on an intercept-only non-mixture model with scale parameter 1.  Standard deviations are chosen so that 95\% of the prior detection probability rests within $p_{sr}^{(det)} \in (0.01, 1.0)$.  Priors for Intercept$^A$ center at a median abundance of about 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).  Because covariates are centered and scaled, priors for effect parameters are centered at zero with standard deviations matching the appropriate intercept term.  For all shape and variance parameters, we choose half-Cauchy(0,1) priors, and for the mixing parameter $\gamma$ we chose a uniform prior.  All priors are independent.
\begin{align*}
\text{Intercept}^D &\sim N(-2.7, 2.25^2) \text{ or } N(-2.3, 1.2^2)\\
\text{Intercept}^A &\sim N(1, 1)\\
\boldsymbol{\beta}^D &\sim N(0, 2.25^2)\\
\boldsymbol{\beta}^A &\sim N(0, 1)\\
\sigma_D, \sigma_A, \alpha, \sigma_{det}, k &\sim \text{half-Cauchy}(0,1)\\
\gamma &\sim \text{Beta}(1,1)
\end{align*}

>>>>

\newcommand{\vn}{\textbf{n}}

@@

\jarad{The introduction needs to end earlier. It should be about 1-2 pages covering what has previously been done and outline our approach. So this should probably start a new section and before this there should be an outline of our manuscript. }

@@

\begin{figure}[h!]\centering
\includegraphics[width=0.68\textwidth]{Hazards.pdf}
\caption{\label{HazardFxns}  Examples of instantaneous rates of detection (hazard functions) from exponential, gamma, lognormal, and Weibull TTDDs. \jarad{Are these setup to have the same mean and variance or anything like that?}
\end{figure}

@@

Our model is formulated for single-visit removal-sampled point-count surveys conducted at a set of sites $s=1,\dotso,S$ repeated across distinct time periods $r=1,\dotso,R$. \jarad{So time periods are R because we already use t for intervals? Argh!}  We present the model for times to detection censored into $i=1,\dotso,I$ intervals \jarad{Why don't we present the exact time-to-detection model first and then censoring it?}, but it is straight-forward to simplify the model for exact time-to-detection data.  
We denote the number of individual observed at site $s$, time $r$, and in interval $i$ as $n_{sri}$ and define the total number of individuals observed during survey as $\vn_{sr} = (\sum_{i=1}^I n_{sri})n_{sr1},\ldots,n_{srI})$.

@@

\jarad{We should try to avoid using subsubsections. Perhaps we should provide a data section and then a model section?}

\subsubsection{Data Model}
We assume that counts from different surveys are conditionally independent.  We also assume that the time-to-detection is independent, identically distributed for all individuals of a species at any given survey.  Given that an organism is detected, we denote the probability it is first detected during interval $i$ as $p_{sri}$ such that $\sum_{i=1}^I p_{sri} = 1$.  We write $\textbf{p}_{sr}$ to denote the vector of interval-specific detection probabilities at any given survey.  The data model is then:
\begin{align}\label{eq:datamodel}
\vn_{sr} &\sim \mbox{Multinomial}\left(n_{sr}^{(obs)}, \textbf{p}_{sr}\right)\\
n_{sr}^{(obs)} &\sim \mbox{Binomial}\left(N_{sr}, p_{sr}^{(det)}\right)
\end{align}

\jarad{Pedogogically, I would introduce the binomial model first and then the multinomial model. Or we could introduce a Multinomial with parameter $N_{sr}$ and an additional bin for non-detection.}

@@

\subsubsection{Detection Model}\label{sec:detectionmodel}

\jarad{I think you should introduce just the exponential model first with no covariates. Thus, relative to what has been done before, you are only adding in the time-to-detection distribution. Later you can introduce other TTDDs and covariates.}

We model $\textbf{p}_{sr}$ and $p_{sr}^{(det)}$ by means of the partitioned cdf of a TTDD, which we express generically as $F_T(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})$. 
\jarad{Do we want to introduce explanatory variables already? Perhaps it would be smoother to wait until later to introduce them.}
Here, $T$ is a random variable for the first time at which an individual is detected, $\textbf{X}_{sr}^D$ and $\textbf{Z}_{sr}^D$ are detection-related covariates and random effect factor levels to be used in a mixed effects linear regression, and $\boldsymbol{\theta}$ is a vector of parameters appropriate to the TTDD being used.  In this study, we consider eight different forms for the TTDD.  The first four TTDDs we employ are:
\jarad{You don't need an enumerate here. Would it be easier to introduce the exponential first and then the other 3 altogether later?}
\begin{enumerate}
\item $T_{srq} \overset{ind}{\sim} \text{Exponential}(\varphi_{sr})$
\item $T_{srq} \overset{ind}{\sim} \text{Gamma}(\alpha, \varphi_{sr})$
\item $T_{srq} \overset{ind}{\sim} \text{Weibull}(k, \varphi_{sr})$
\item $T_{srq} \overset{ind}{\sim} \text{Lognormal}(\mu_{sr}, \sigma_{det}^2)$
\end{enumerate}
where $q = 1,\dotso,N_{sr}$ indexes individuals present at survey $sr$, and $\varphi_{sr}$ is always a survey-specific detection rate parameter.  We treat the shape parameters $\alpha$, $\sigma_{det}^2$, and $k$ as constant across all surveys.

@@

For exponential, gamma, and Weibull TTDDs, we model the log of the rate parameter $\varphi_{sr}$ via mixed effects regression using survey-specific covariates.  In so doing, we are equivalently modeling the log of the expected time-to-detection, which is inversely related to the detection rate.  For the lognormal distribution, we fit a regression to $-\mu_{sr}$.  Thus, the link functions in our detection model are:
\begin{alignat}{3}
&\text{Exponential:\;} &&\log(\varphi_{sr}) &&= \text{Intercept}^D + \textbf{X}_{sr}^D\boldsymbol{\beta}^D + \textbf{Z}_{sr}^D\boldsymbol{\xi}^D = -\log(E[T_{srq}])\\
&\text{Gamma:} &&\log(\varphi_{sr}) &&= \text{Intercept}^D + \textbf{X}_{sr}^D\boldsymbol{\beta}^D + \textbf{Z}_{sr}^D\boldsymbol{\xi}^D = -\log(E[T_{srq}]) + \log(\alpha)\\
&\text{Weibull:}  &&\log(\varphi_{sr}) &&= \text{Intercept}^D + \textbf{X}_{sr}^D\boldsymbol{\beta}^D + \textbf{Z}_{sr}^D\boldsymbol{\xi}^D = -\log(E[T_{srq}]) + \log\left(\Gamma\left(1 + 1/k\right)\right)\\
&\text{LogNormal:} &&-\mu_{sr} &&= \text{Intercept}^D + \textbf{X}_{sr}^D\boldsymbol{\beta}^D + \textbf{Z}_{sr}^D\boldsymbol{\xi}^D = -\log(E[T_{srq}]) + \sigma_{det}^2/2
\end{alignat}
\jarad{Are these necessary?}
where $\boldsymbol{\beta}^D$ is a vector of fixed effects, and $\xi_i^D \overset{ind}{\sim} N(0,\sigma_{D[i]}^2)$ are random effects, with $\sigma_{D[i]}^2$ denoting the variance associated with $\xi_i^D$.  When the TTDD is a mixture distribution, we model $F_{T1}$ portion of the mixture according to the above link functions and treat the mixing parameter $\gamma$ as constant across all surveys.

For interval-censored data, we denote the cutoffs for each observation interval with $C_i$, where $C_0 = 0 \text{ and } C_I = \text{the survey duration}$.  Conditioned on an individual being detected \jarad{Would this be easier to introduce without conditioning on individuals being detected?}, we calculate interval-specific detection probabilities $p_{sri}$ by partitioning the cdf of the TTDD at the observation interval cutoffs and normalizing by the binomial detection probability $p_{sr}^{(det)}$.  In summary, the detection model is:
\begin{align}
p_{sri} &= \left[F_T(C_i|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta}) - F_T(C_{i-1}|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})\right]\big/p_{sr}^{(det)}\\
\label{eq:pdet}p_{sr}^{(det)} &= F_T(C_I|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})\\
F_T(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta}) &= \text{a TTDD as described above}
\end{align}

To adapt our model for exact time-to-detection data, we only need to replace the multinomial distribution in Eq. (\ref{eq:datamodel}).  Instead of modeling the interval-specific counts $\vn_{sr}$, we model the times to detection directly.  Defining $\textbf{T}_{sr}$ as a vector of all observed detection times for a given survey, and continuing to treat all detections as independent, we write
\[
Pr\left(\textbf{T}_{sr}|n_{sr}^{(obs)}\right) = \prod\limits_{q=1}^{n_{sr}^{(obs)}} f_T(t_{srq}|t_{srq}<C_I)
\]
The rest of the model remains the same, except there is no need to calculate interval-specific times to detection.  Partial finite-mixture TTDDs clearly only make sense in the interval-censored case.

\subsubsection{Abundance Model}

\jarad{Again, I would introduce this without covariates or random effects first.

For simplicity, we model abundance at each site during each year $N_{sr}$ using a Poisson distribution with a survey-specific expected value of $\lambda_{sr}$, though any non-negative integer-valued distribution may be used.  We can decompose $N_{sr}$ into observed and unobserved portions.  The unobserved portion follows a binomial-Poisson hierarchy, resulting in Equation \ref{eq:binPois}.  We model expected abundance $\lambda_{sr}$ using a mixed effects linear regression with a log link.

@@

\jarad{Put a discussion of priors in a new section. You should introduce the general form of the priors with specific values and specify the specific values in the simulation study and real data analysis sections.}

@@

\jarad{You need to include a section about the MCMC, e.g. Stan/JAGS how to monitor convergence, etc.}

@@

\jarad{This should not be a model (or methods section). I think you should put a Data section earlier that introduces the data collection methods. Then the rest of this section should go much later, after the simulation study, in a real data analysis section. }

@@

\subsubsection{Priors}\label{sec:priors}
Our goal in selecting priors is to be non-informative within a reasonable range of values.  Priors for Intercept$^D$ are centered so that the median detection probability is $p_{sr}^{(det)} = 0.50$, based on an intercept-only non-mixture model with scale parameter 1.  Standard deviations are chosen so that 95\% of the prior detection probability rests within $p_{sr}^{(det)} \in (0.01, 1.0)$.  Priors for Intercept$^A$ center at a median abundance of about 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).  Because covariates are centered and scaled, priors for effect parameters are centered at zero with standard deviations matching the appropriate intercept term.  For all shape and variance parameters, we choose half-Cauchy(0,1) priors, and for the mixing parameter $\gamma$ we chose a uniform prior.  All priors are independent.
\jarad{This takes up too much space.}
\begin{align*}
\text{Intercept}^D &\sim N(-2.7, 2.25^2) \text{ or } N(-2.3, 1.2^2)\\
\text{Intercept}^A &\sim N(1, 1)\\
\boldsymbol{\beta}^D &\sim N(0, 2.25^2)\\
\boldsymbol{\beta}^A &\sim N(0, 1)\\
\sigma_D, \sigma_A, \alpha, \sigma_{det}, k &\sim \text{half-Cauchy}(0,1)\\
\gamma &\sim \text{Beta}(1,1)
\end{align*}

\jarad{New section for simulation studies with the results.}
